---
title: "Potential Daily Nexus"
output: html_document
---

# Trying to see if its true that a proffsseor that teachs a course over a long period of time has better results measured with the average gpa with said course

<!-- Needed Librbaies -->

``` {r}
library(dplyr)
library(tidyr)
```

<!-- Variables and Dataframe --->

  ``` {r}
grades <- read.csv("allGrades.csv")
startYear <- 2020
endYear <- 2023
selectQuarter <- "Spring"
  ```

<!-- Filtering the data by year and quarter -->

``` {r}
grades <- grades %>%
  filter(as.numeric(year) >= startYear & as.numeric(year) <= endYear) %>%
  filter(quarter == selectQuarter)

```


<!-- Creating the base of two new tables -->

``` {r} 
baseTable <- data.frame(instructors = grades$instructor, course = grades$course, dept =  grades$dept, quarter = grades$quarter, year = grades$year)
```

 <!-- Calculating the pass fail rate of each course and adding it to a new table based off of baseTable -->
```{r}
process_quarter <- function(grades, quarter, output_file) {
  # Filter data by quarter
  quarter_grades <- grades %>%
    filter(as.numeric(year) >= startYear & as.numeric(year) <= endYear) %>%
    filter(quarter == quarter)
  
  # Create base table
  baseTable <- data.frame(
    instructors = quarter_grades$instructor, 
    course = quarter_grades$course, 
    dept = quarter_grades$dept, 
    quarter = quarter_grades$quarter, 
    year = quarter_grades$year
  )
  
  # Calculate passFailRatio
  passNoTable <- baseTable
  passNoTable$passFailRatio <- NA
  
  for (row in 1:nrow(quarter_grades)) {
    currentRow <- quarter_grades[row, ]
    amtPassed <- sum(currentRow$Ap, currentRow$Bp, currentRow$Cp, currentRow$Am, currentRow$Bm, currentRow$P, na.rm = TRUE)
    amtFailed <- sum(currentRow$Cm, currentRow$Dm, currentRow$C, currentRow$D, currentRow$F, na.rm = TRUE)
    sumOfPassAndFail <- amtPassed + amtFailed
    
    if (sumOfPassAndFail > 0) {
      passNoTable$passFailRatio[row] <- amtPassed / sumOfPassAndFail
    } else {
      passNoTable$passFailRatio[row] <- NA
    }
  }
  
  # Filter out NA rows
  passNoTable <- passNoTable %>% filter(!is.na(passFailRatio))
  
  # Filter for instructors teaching the same course for the entire period
  trendTable <- passNoTable %>%
    group_by(instructors, course, dept) %>%
    filter(n_distinct(year) == (endYear - startYear + 1)) %>%
    ungroup()
  
  # Reshape to wide format
  trendTableWide <- trendTable %>%
    select(instructors, course, dept, year, passFailRatio) %>%
    pivot_wider(names_from = year, values_from = passFailRatio, names_prefix = "passFailRatio")
  
  # Define the Trend calculation function
  calculate_trend <- function(ratios) {
    tally <- 0
    for (i in seq_along(ratios)[-length(ratios)]) {
      if (!is.na(ratios[[i]]) && !is.na(ratios[[i + 1]])) {
        if (ratios[[i]] > ratios[[i + 1]]) {
          tally <- tally - 1
        } else {
          tally <- tally + 1
        }
      }
    }
    if (tally <= -3) {
      return("Strong Downward")
    } else if (tally == -2) {
      return("Moderate Downward")
    } else if (tally == -1) {
      return("Slight Downward")
    } else if (tally == 0) {
      return("No Trend")
    } else if (tally == 1) {
      return("Slight Upward")
    } else if (tally == 2) {
      return("Moderate Upward")
    } else {
      return("Strong Upward")
    }
  }
  
  # Add Trend column
  trendTableWide <- trendTableWide %>%
    rowwise() %>%
    mutate(Trend = calculate_trend(c_across(starts_with("passFailRatio")))) %>%
    ungroup()
  
  # Write to CSV
  write.csv(trendTableWide, output_file, row.names = FALSE)
}

```

<!-- Testing Area -->

``` {r}
# Apply the function for each quarter
process_quarter(grades, "Winter", "final_winter.csv")
process_quarter(grades, "Fall", "final_fall.csv")
process_quarter(grades, "Spring", "final_spring.csv")
```

Do profs who teach the same class over a long period of time teach better than they did at the start?

Measure?
avgGPA -> studenside

as a student I want a good grade so if a prof is grading harsher over time I think they are teaching worse

Passing to not passing rate   

as a prof I don't care about the grades of students I care whether or not they understand the material I am teaching them


instead of all departments make it so that it is the top ten departments
 
 student inetrviews
 department chair aS well



